---
title: "Qari_Final_Project_V2.Rmd"
author: "Mubashira Qari"
date: "April 30, 2022"
output: html_document
---

### Overview: 

Pymaceuticals Inc., a fictional burgeoning pharmaceutical company based out of San Diego, CA, specializes in drug-based, anti-cancer pharmaceuticals.They have provided the data to test the efficacy of potential drug treatments for squamous cell carcinoma. In this study, 249 mice identified with Squamous cell carcinoma (SCC) tumor growth, kind of skin cancer, were treated through a variety of drug regimens. Over the course of 45 days, tumor development was observed and measured.The objective is to analyze the data to show how four treatments (Capomulin, Infubinol, Ketapril, and Placebo) compare.


### Research question:
### You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.

Question 1: Is Capomulin more or less effective in reducing the tumor size than Infubinol, or Ketapril drugs categories? 

Question 2: Is there a correlation between the age and the tumor size growth for each drug category?

Question 3: Is there a correlation between the weight and the tumor size growth for each drug category?

Hypothesis Test:

Null Hypothesis: There is no difference between the mean percent change in tumor volume for the four drug categories.

Alternate Hypothesis: There is a difference between the mean percent change in tumor volume for the four drug categories.

Approach for answering the research question will be: 

1- Calculate the mean percent change in tumor volume for the four drug categories.

2- Perform the Hypothesis test to find out whether or not the difference exist between the mean tumor size for all four drug categories.

3- Perform linear regression to study the correlation between various variables by calculating the correlation coefficient.

4- Finally analyze the results to find out if Capomulin more or less effective in reducing the tumor size of sample mice than Infubinol, or Ketapril drugs categories.


### Cases:
### What are the cases? How many different drug treatments are there? How many total sample size as well as the sample size by drug treatments are there?

Answer: The metadata_df contain 249 unique mouse id and so are the number of cases that treated with variety of drug regimem .The results_df dataset holds the tumor growth measurments observed for each Mouse ID and carries 1,893 rows results. There are 10 different drug treatments. The total sample size of mouse_id for four treatments (Capomulin, Infubinol, Ketapril, and Placebo) is 100 and the sample size of mouse_id by drug treatments is 25 each.


### Data collection:
### Describe the method of data collection.

Answer: Data is collected by the fictitious pharmaceutical company who was testing the efficacy of potential drug treatments for squamous cell carcinoma.
I import the data into my .Rmd file from github.


### Type of study:
### What type of study is this (observational/experiment)?

Answer: This is an experimental study.A group of 249 mice were monitored after administration of a variety of drug regimens over a 45-day treatment period. The impact of Capomulin on tumor growth, metastasis and survival rates were monitored, along with Infubinol, Ketapril, and Placebo.


### Data Source:
### If you collected the data, state self-collected. If not, provide a citation/link.

Answer: The citation and data collection links are as follows. 

In my search for the experimental datasets, I found the Mouse_metadata and the Study_results on the GitHub link provided below:

https://raw.githubusercontent.com/rfpoulos/pymaceuticals/master/data/Mouse_metadata.csv

https://raw.githubusercontent.com/rfpoulos/pymaceuticals/master/data/Study_results.csv

Upon further research in finding the original source of the the dataset, I found that these datasets are provided by Pymaceuticals Inc., a fictional burgeoning pharmaceutical company based out of San Diego, CA, specializes in drug-based, anti-cancer pharmaceuticals. Below is the link for the original source of the datasets.


https://c-l-nguyen.github.io/web-design-challenge/index.html



### Response
### What is the response variable, and what type is it (numerical/categorical)?

Answer: The response variable is the size of tumor, "Tumor.Volume..mm3." and it holds a numerical data.

### Explanatory

What is the explanatory variable, and what type is it (numerical/categorical)?

Answer: The explanatory variable is the "Drug.Regimen" and it holds a categorical data and "Timepoint" which holds numerical data. The 'Timepoint' unit is 'days'.

### Relevant summary statistics: (Tables and Charts)

Provide summary statistics relevant to your research question. For example, if you’re comparing means across groups provide means, SDs, sample sizes of each group. This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.

### Libraries Imported

```{r}
library(tidyverse)
library(dplyr)
library(plotly)
library(tidyr)
library(stringr)
library(psych)
library(ggplot2)
```

### Data Preparation

#### load dataset1

```{r }
metadata_df <- read.delim("https://raw.githubusercontent.com/rfpoulos/pymaceuticals/master/data/Mouse_metadata.csv", header=T, sep=",")
head(metadata_df)
```

### Grouping by Drug.Regimen

```{r}
df <- metadata_df %>%
  group_by(Drug.Regimen) 

head(df)
```

### Load dataset2

```{r}
results_df <- read.delim("https://raw.githubusercontent.com/rfpoulos/pymaceuticals/master/data/Study_results.csv", header=T, sep=",")
head(results_df)
```


```{r}
summary(metadata_df)
```

### Summary Statistic

```{r}
summary(results_df)
```

### Sample Sizes for metadata_df

```{r}
nrow(metadata_df)
```

### Sample Sizes for results_df

```{r}
nrow(results_df)
```

### How many drug treatments are there?

```{r}

drug_count <- unique(metadata_df$Drug.Regimen)

drug_count
```

```{r}

length(drug_count)
```

### Sample sizes of mouse_id by drug treatment

```{r}
capomulin_df <- filter(metadata_df, Drug.Regimen=="Capomulin")

head(capomulin_df)
```

```{r}
nrow(capomulin_df)
```

```{r}
infubinol_df <- filter(metadata_df, Drug.Regimen=="Infubinol")

nrow(infubinol_df)
```

```{r}
ketapril_df <- filter(metadata_df, Drug.Regimen=="Ketapril")

nrow(ketapril_df)
```

```{r}
placebo_df <- filter(metadata_df, Drug.Regimen=="Placebo")

nrow(placebo_df)
```

### Performing full outer join, so that no data is lost

```{r}
merge_df <- merge(x = metadata_df, y = results_df, all = TRUE)

head(merge_df)
```

```{r}
glimpse(merge_df)
```

### Dropping the NA rows

```{r}
merge_df <- merge_df %>% drop_na()

head(merge_df)
```


### Change colnames of some columns
### assigning new names to the columns of the merged data frame
### Colnames(df)[2] <- "new_col2"

```{r}
colnames(merge_df)[1] <- c("Mouse_Id")
colnames(merge_df)[2] <- c("Drug_Regimen")
colnames(merge_df)[5] <- c("Weight_g")
colnames(merge_df)[7] <- c("Tumor_Volume_mm3")
colnames(merge_df)[8] <- c("Metastatic_Sites")

head(merge_df)
```

```{r}
merge_df %>% group_by(Mouse_Id, Timepoint)

head(merge_df)
```


### Remove duplicate rows across entire data frame

```{r}

merge_df <- merge_df[!duplicated(merge_df), ]

head(merge_df)
```

### filter by Capomulin, Infubinol, Ketapril, and Placebo

```{r}

capomulin_df <- filter(merge_df, Drug_Regimen == "Capomulin")
infubinol_df <- filter(merge_df, Drug_Regimen == "Infubinol")
ketapril_df <- filter(merge_df, Drug_Regimen == "Ketapril")
placebo_df <- filter(merge_df, Drug_Regimen == "Placebo")


head(capomulin_df)
```

### The initial Tumor Volume for starting point zero is 45 mm3 for all mice
### capomulin_df

```{r}
# Filtering the max timepoint values

capo_df <- select(capomulin_df, Mouse_Id, Drug_Regimen, Timepoint, Tumor_Volume_mm3, Age_months, Weight_g) %>%
   group_by(Mouse_Id) %>%
  filter(Timepoint == max(Timepoint, na.rm=TRUE))


head(capo_df)
```



```{r}
# Calculating the percent change in Tumor volume
capo_df <- capo_df %>% 
  mutate(percent_change_Tumor_Volume = ((Tumor_Volume_mm3-45)/45)*100)
head(capo_df)
```

### The initial Tumor Volume for starting point zero is 45 mm3 for all mice
### Infubinol_df

```{r}
# Filtering the max timepoint values
infu_df <- select(infubinol_df, Mouse_Id, Drug_Regimen, Timepoint, Tumor_Volume_mm3, Age_months, Weight_g) %>%
   group_by(Mouse_Id) %>%
  filter(Timepoint == max(Timepoint, na.rm=TRUE))

# Calculating the percent change in Tumor volume
infu_df <- infu_df %>% 
  mutate(percent_change_Tumor_Volume = ((Tumor_Volume_mm3-45)/45)*100)

head(infu_df)
```

### The initial Tumor Volume for starting point zero is 45 mm3 for all mice
### Ketapril_df

```{r}
# Filtering the max timepoint values
keta_df <- select(ketapril_df, Mouse_Id, Drug_Regimen, Timepoint, Tumor_Volume_mm3, Age_months, Weight_g) %>%
   group_by(Mouse_Id) %>%
  filter(Timepoint == max(Timepoint, na.rm=TRUE))

# Calculating the percent change in Tumor volume
keta_df <- keta_df %>% 
  mutate(percent_change_Tumor_Volume = ((Tumor_Volume_mm3-45)/45)*100)

head(keta_df)
```

### The initial Tumor Volume for starting point zero is 45 mm3 for all mice
### Placebo_df

```{r}
plac_df <- select(placebo_df, Mouse_Id, Drug_Regimen, Timepoint, Tumor_Volume_mm3, Age_months, Weight_g) %>%
   group_by(Mouse_Id) %>%
  filter(Timepoint == max(Timepoint, na.rm=TRUE))

# Calculating the percent change in Tumor volume
plac_df <- plac_df %>% 
  mutate(percent_change_Tumor_Volume = ((Tumor_Volume_mm3-45)/45))

head(plac_df)
```


```{r}

# put all data frames into list
df_list <- list(capo_df, infu_df, keta_df, plac_df)      

#merge all data frames together
new_df <- Reduce(function(x, y) merge(x, y, all=TRUE), df_list)  
head(new_df)
```

```{r}
new_df$percent_change_Tumor_Volume = round(new_df$percent_change_Tumor_Volume, digit=2)
head(new_df)
```



```{r}
ggplot(new_df, aes(x=Drug_Regimen, y=percent_change_Tumor_Volume, fill=Drug_Regimen)) + 
       geom_bar(stat="identity") +
       coord_flip()
```




### Assumptions
The ANOVA test makes the following assumptions about the data:

Independence of the observations. Each subject should belong to only one group. There is no relationship between the observations in each group. Having repeated measures for the same participants is not allowed.
No significant outliers in any cell of the design
Normality. the data for each design cell should be approximately normally distributed.
Homogeneity of variances. The variance of the outcome variable should be equal in every cell of the design.

```{r}
test_df <- select(new_df, Drug_Regimen, percent_change_Tumor_Volume) %>%
  group_by(Drug_Regimen) %>%
 summarise(percent_change_Tumor_Volume = mean(percent_change_Tumor_Volume, na.rm=TRUE))

head(test_df)
```

### Summary statistics

Compute some summary statistics (count, mean and sd) of the variable weight organized by groups:

```{r}
new_df %>%
   group_by(Drug_Regimen) %>%
  get_summary_stats(percent_change_Tumor_Volume, type = "mean_sd")
```
### Visualization

Create a box plot of percent_change_Tumor_Volume by Drug_Regimen:

```{r}
ggboxplot(new_df, x = "Drug_Regimen", y = "percent_change_Tumor_Volume")
```

### Check assumptions
### Outliers

```{r}
new_df %>% 
  group_by(Drug_Regimen) %>%
  identify_outliers(percent_change_Tumor_Volume)
```
In this situation there is one outliers, this can be due to data entry errors, measurement errors or unusual values and the mean is not effected.

### Normality assumption

Check normality assumption by analyzing the model residuals. 
QQ plot and Shapiro-Wilk test of normality are used.

### QQ plot draws the correlation between a given data and the normal distribution.

```{r}
# Build the linear model
model  <- lm(percent_change_Tumor_Volume ~ Drug_Regimen, data = new_df)
# Create a QQ plot of residuals
ggqqplot(residuals(model))
```

### Check normality assumption by groups

Computing Shapiro-Wilk test for each group level. If the data is normally distributed, the p-value should be greater than 0.05.
```{r}
new_df %>%
  group_by(Drug_Regimen) %>%
  shapiro_test(percent_change_Tumor_Volume)
```

The score were normally distributed (p > 0.05) for each group, as assessed by Shapiro-Wilk’s test of normality.

### Create QQ plots for each group level:

QQ plot draws the correlation between a given data and the normal distribution. 

```{r}
ggqqplot(new_df, "percent_change_Tumor_Volume", facet.by = "Drug_Regimen")
```

All the points fall approximately along the reference line, for each cell. So we can assume normality of the data.

### Homogneity of variance assumption

The residuals versus fits plot can be used to check the homogeneity of variances.

```{r}
plot(model, 1)
```

In the plot above, there is no evident relationships between residuals and fitted values (the mean of each groups), which is good. So, we can assume the homogeneity of variances.





#========================================================

```{r}
head(capo_df)
```

```{r}
#fit regression model
model <- lm(Tumor_Volume_mm3~Weight_g+Age_months, data=capo_df)

#view model summary
summary(model)

#By looking at the coefficients, it seems as though weight is accounting for most of the variance (B = 1.76, p<.05). Age is not a significant predictor of tumor size (B = .053, p = .481).
```

Residuals:
    Min      1Q  Median      3Q     Max 
-7.2433 -0.7903  0.1667  1.5253  4.8761 

This section displays a summary of the distribution of residuals from the regression model. A residual is the difference between the observed value and the predicted value from the regression model.

The minimum residual was -7.2433, the median residual was 0.1667 and the max residual was 4.8761.

Coefficients:
            Estimate Std. Error t value Pr(>|t|)    
(Intercept)  0.48591    4.32578   0.112    0.912    
Weight_g     1.76129    0.20298   8.677 1.49e-08 ***
Age_months   0.05303    0.07401   0.716    0.481    
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

This section displays the estimated coefficients of the regression model. We can use these coefficients to form the following estimated regression equation:

Tumor_Volume_mm3 = 0.48591 + 1.76129*Weight_g + 0.05303*Age_months


