---
title: "Data 606 - Final Project"
output:
  pdf_document: default
  html_document: default
---

---
title: "DATA 606 Data Project Proposal"
output: html_document
---



### Libraries Imported

```{r}
library(tidyverse)
library(dplyr)
library(plotly)
library(tidyr)
library(stringr)
library(psych)
library(ggplot2)
```

### Data Preparation

#### load dataset1

```{r }
metadata_df <- read.delim("https://raw.githubusercontent.com/rfpoulos/pymaceuticals/master/data/Mouse_metadata.csv", header=T, sep=",")
head(metadata_df)
```
### Grouping by Drug.Regimen

```{r}
df <- metadata_df %>%
  group_by(Drug.Regimen) 

head(df)
```
### Load dataset2

```{r}
results_df <- read.delim("https://raw.githubusercontent.com/rfpoulos/pymaceuticals/master/data/Study_results.csv", header=T, sep=",")
head(results_df)
```

Introduction: Pymaceuticals Inc., a fictional burgeoning pharmaceutical company based out of San Diego, CA, specializes in drug-based, anti-cancer pharmaceuticals.They have provided the data to test the efficacy of potential drug treatments for squamous cell carcinoma. In this study, 249 mice identified with Squamous cell carcinoma (SCC) tumor growth, kind of skin cancer, were treated through a variety of drug regimens. Over the course of 45 days, tumor development was observed and measured.The objective is to analyze the data to show how four treatments (Capomulin, Infubinol, Ketapril, and Placebo) compare.


### Research question:
### You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.

Question 1: Is Capomulin more effective than the three other drugs in the dataset? 

Question 2: Is there a correlation between the age, weight and the effectiveness of capomulin?

Null Hyothesis: There is no difference between the effectiveness of the four drug regimens.

Alternate Hyothesis: Capomulin is more effective than the other three drug regimens on treating SCC tumor growth.


Approach for answering the research question will be: 


1- Perform linear regression to study the correlation between various variables and calculating the correlation coefficient.

2- And finally compare the four population against each other.

3- Perform Hypothesis testing 

### Cases:
### What are the cases? How many different drug treatments are there? How many total sample size as well as the sample size by drug treatments are there?

Answer: The metadata_df contain 249 unique mouse id and so are the number of cases that treated with variety of drug regimem .The results_df dataset holds the tumor growth measurments observed for each Mouse ID and carries 1,893 rows results. There are 10 different drug treatments. The total sample size of mouse_id for four treatments (Capomulin, Infubinol, Ketapril, and Placebo) is 100 and the sample size of mouse_id by drug treatments is 25 each.


### Data collection:
### Describe the method of data collection.

Answer: Data is collected by the fictitious pharmaceutical company who was testing the efficacy of potential drug treatments for squamous cell carcinoma.
I import the data into my .Rmd file from github.


### Type of study:
### What type of study is this (observational/experiment)?

Answer: This is a experimental study.A group of 249 mice were monitored after administration of a variety of drug regimens over a 45-day treatment period. The impact of Capomulin on tumor growth, metastasis and survival rates were monitored, along with Infubinol, Ketapril, and Placebo.


### Data Source:
### If you collected the data, state self-collected. If not, provide a citation/link.

Answer: The citation and data collection links are as follows. 

In my search for the experimental datasets, I found the Mouse_metadata and the Study_results on the GitHub link provided below:

https://raw.githubusercontent.com/rfpoulos/pymaceuticals/master/data/Mouse_metadata.csv

https://raw.githubusercontent.com/rfpoulos/pymaceuticals/master/data/Study_results.csv

Upon further research in finding the original source of the the dataset, I found that these datasets are provided by Pymaceuticals Inc., a fictional burgeoning pharmaceutical company based out of San Diego, CA, specializes in drug-based, anti-cancer pharmaceuticals. Below is the link for the original source of the datasets.


https://c-l-nguyen.github.io/web-design-challenge/index.html



### Response
### What is the response variable, and what type is it (numerical/categorical)?

Answer: The response variable is the size of tumor, "Tumor.Volume..mm3." and it holds a numerical data.

### Explanatory

What is the explanatory variable, and what type is it (numerical/categorical)?

Answer: The explanatory variable is the "Drug.Regimen" and it holds a categorical data and "Timepoint" which holds numerical data. The 'Timepoint' unit is 'days'.

### Relevant summary statistics: (Tables and Charts)

Provide summary statistics relevant to your research question. For example, if you’re comparing means across groups provide means, SDs, sample sizes of each group. This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.

```{r}
summary(metadata_df)
```

### Summary Statistic

```{r}
summary(results_df)
```

### Sample Sizes for metadata_df

```{r}
nrow(metadata_df)
```

### Sample Sizes for results_df

```{r}
nrow(results_df)
```

### How many drug treatments are there?

```{r}

drug_count <- unique(metadata_df$Drug.Regimen)

drug_count
```

```{r}

length(drug_count)
```

### Sample sizes of mouse_id by drug treatment

```{r}
capomulin_df <- filter(metadata_df, Drug.Regimen=="Capomulin")

head(capomulin_df)
```

```{r}
nrow(capomulin_df)
```

```{r}
infubinol_df <- filter(metadata_df, Drug.Regimen=="Infubinol")

nrow(infubinol_df)
```

```{r}
ketapril_df <- filter(metadata_df, Drug.Regimen=="Ketapril")

nrow(ketapril_df)
```

```{r}
placebo_df <- filter(metadata_df, Drug.Regimen=="Placebo")

nrow(placebo_df)
```

### Performing full outer join, so that no data is lost

```{r}
merge_df <- merge(x = metadata_df, y = results_df, all = TRUE)

head(merge_df)
```

```{r}
glimpse(merge_df)
```

### Dropping the NA rows

```{r}
merge_df <- merge_df %>% drop_na()

head(merge_df)
```


### Change colnames of some columns
### assigning new names to the columns of the merged data frame
### Colnames(df)[2] <- "new_col2"

```{r}
colnames(merge_df)[1] <- c("Mouse_Id")
colnames(merge_df)[2] <- c("Drug_Regimen")
colnames(merge_df)[5] <- c("Weight_g")
colnames(merge_df)[7] <- c("Tumor_Volume_mm3")
colnames(merge_df)[8] <- c("Metastatic_Sites")

head(merge_df)
```

```{r}
merge_df %>% group_by(Mouse_Id, Timepoint)

head(merge_df)
```


```{r}
df1 <- select(merge_df, Drug_Regimen, Tumor_Volume_mm3, Age_months, Weight_g)
head(df1)
```

```{r}
df1 <- group_by(df1, Drug_Regimen)
head(df1)
```

### Finding the summary statistics of Tumor_Volume

```{r}
stats_df <- df1 %>% summarise(
  Tumor_Volume_mean = mean(Tumor_Volume_mm3), Tumor_Volume_median = median(Tumor_Volume_mm3), Tumor_Volume_sd = sd(Tumor_Volume_mm3), Tumor_Volume_se = sd(Tumor_Volume_mm3)/sqrt(length((Tumor_Volume_mm3))))

head(stats_df)
```

### Comparing means of tumor size by drug treatment.

```{r}
library(ggplot2)

# plot mean salaries
ggplot(stats_df, 
       aes(x = Drug_Regimen, 
           y = Tumor_Volume_mean)) +
  geom_bar(stat = "identity",  fill = "cornflowerblue")
```

### Side-by-side box plots are very useful for comparing groups (i.e., the levels of a categorical variable) on a numerical variable. Outliers are prominent for Drug_Regimen Capomulin, Propriva, Ramicane and Stelasyn.  

```{r}
ggplot(merge_df, 
       aes(x = Drug_Regimen, 
           y = Tumor_Volume_mm3)) +
  geom_boxplot() +
  labs(title = "Mean distribution by Drug_Regimen")

```

### Finding the mice count of each Drug Regimen

```{r}
count_df <- df1 %>% count(Drug_Regimen)

count_df
```

### Ploting the number of mice in each drug regimen

```{r}
barplot(c(230, 178, 178, 188, 186, 181, 161, 228, 181, 182),
        names.arg=c("Capomulin","Ceftamin","Infubinol","Ketapril","Naftisol", "Placebo", "Propriva", "Ramicane", "Stelasyn", "Zoniferol"),
        ylim=c(0,250),
        col=c("beige","orange","lightgreen","lightblue","yellow", "blue", "green", "pink", "purple", "red"),
        ylab="Count of Mice per Drug Regimen")
```


### Remove duplicate rows across entire data frame

```{r}

merge_df <- merge_df[!duplicated(merge_df), ]

head(merge_df)
```

### filter by Capomulin, Infubinol, Ketapril, and Placebo

```{r}

capomulin_df <- filter(merge_df, Drug_Regimen == "Capomulin")
infubinol_df <- filter(merge_df, Drug_Regimen == "Infubinol")
ketapril_df <- filter(merge_df, Drug_Regimen == "Ketapril")
placebo_df <- filter(merge_df, Drug_Regimen == "Placebo")


head(capomulin_df)
```

### To generate a scatter plot of average tumor volume vs. mouse weight for all mice in the Capomulin regimen.
### First we calculate the final tumor volume of each mouse_id across four of the treatment regimens:  
### (Capomulin, Infubinol, Ketapril, and Placebo)

### Since not all mice lived until timepoint 45, we start by getting the last (greatest) timepoint for each mouse

### capomulin_df:

```{r}
capo_df1 <- select(capomulin_df, Mouse_Id, Timepoint, Tumor_Volume_mm3) %>%
  group_by(Mouse_Id) %>%
  filter(Timepoint == max(Timepoint, na.rm=TRUE))

head(capo_df1)
```

### Find the average weight by mice_id in Capomulin_df

```{r}
capo_df2 <- select(capomulin_df, Mouse_Id, Weight_g) %>%
  group_by(Mouse_Id) %>%
 summarise(Average_weight = mean(Weight_g, na.rm=TRUE))

head(capo_df2)
```

### Joining the two df's for adding average weight

```{r}

capo_df <- capo_df1 %>% inner_join(capo_df2, by = "Mouse_Id")

head(capo_df)
```

### Find the average age by mice_id in Capomulin_df

```{r}
capo_df3 <- select(capomulin_df, Mouse_Id, Age_months) %>%
  group_by(Mouse_Id) %>%
 summarise(Average_age = mean(Age_months, na.rm=TRUE))

head(capo_df3)
```

### Joining the two df's for adding average age

```{r}

capo_df <- capo_df %>% inner_join(capo_df3, by = "Mouse_Id")

head(capo_df)
```

### summerize the Tumor_Volume_mm3

```{r}

capo_df$Tumor_Volume_mm3 %>%
  summary()
```

### Standard Deviation

```{r}
capo_df$Tumor_Volume_mm3 %>% sd()
```


### For project proposal, plotting correlation matrices with all the relevant variables for Capomulin drug to analyze. 

### capomulin_df Vs Age_months

```{r}

# Creating the plot
plot(capo_df$Average_age, capo_df$Tumor_Volume_mm3, pch = 19, col = "blue")

# Regression line
abline(lm(capo_df$Tumor_Volume_mm3 ~ capo_df$Average_age), col = "red", lwd = 3)

# Pearson correlation
text(paste("Correlation:", round(cor(capo_df$Average_age, capo_df$Tumor_Volume_mm3), 2)), x = 25, y = 95)
```

### capomulin_df Vs Weight_g

```{r}

# Creating the plot
plot(capo_df$Average_weight, capo_df$Tumor_Volume_mm3, pch = 19, col = "blue")

# Regression line
abline(lm(capo_df$Tumor_Volume_mm3 ~ capo_df$Average_weight), col = "red", lwd = 3)

# Pearson correlation
text(paste("Correlation:", round(cor(capo_df$Average_weight, capo_df$Tumor_Volume_mm3), 2)), x = 25, y = 95)
```

### Correlation Matrix

```{r}
pairs(capo_df[,2:5], pch = 19, col = "blue")
```

### Infubinol_df:


```{r}

infu_df1 <- select(infubinol_df, Mouse_Id, Timepoint, Tumor_Volume_mm3) %>%
  group_by(Mouse_Id) %>%
  filter(Timepoint == max(Timepoint, na.rm=TRUE))

### Find the average weight by mice_id in Infubinol_df


infu_df2 <- select(infubinol_df, Mouse_Id, Weight_g) %>%
  group_by(Mouse_Id) %>%
 summarise(Average_weight = mean(Weight_g, na.rm=TRUE))

### Joining the two df's for adding average weight

infu_df <- infu_df1 %>% inner_join(infu_df2, by = "Mouse_Id")

### Find the average age by mice_id in Capomulin_df

infu_df3 <- select(infubinol_df, Mouse_Id, Age_months) %>%
  group_by(Mouse_Id) %>%
 summarise(Average_age = mean(Age_months, na.rm=TRUE))

### Joining the two df's for adding average age

infu_df <- infu_df %>% inner_join(infu_df3, by = "Mouse_Id")

head(infu_df)
```

### summerize the Tumor_Volume_mm3

```{r}

infu_df$Tumor_Volume_mm3 %>%
  summary()
```

### Standard Deviation

```{r}
infu_df$Tumor_Volume_mm3 %>% sd()
```


### infubinol_df Vs Age_months

```{r}

# Creating the plot
plot(infu_df$Average_age, infu_df$Tumor_Volume_mm3, pch = 19, col = "green")

# Regression line
abline(lm(infu_df$Tumor_Volume_mm3 ~ infu_df$Average_age), col = "red", lwd = 3)

# Pearson correlation
text(paste("Correlation:", round(cor(infu_df$Average_age, infu_df$Tumor_Volume_mm3), 2)), x = 25, y = 95)
```

### infubinol_df Vs Weight_g

```{r}

# Creating the plot
plot(infu_df$Average_weight, infu_df$Tumor_Volume_mm3, pch = 19, col = "green")

# Regression line
abline(lm(infu_df$Tumor_Volume_mm3 ~ infu_df$Average_weight), col = "red", lwd = 3)

# Pearson correlation
text(paste("Correlation:", round(cor(infu_df$Average_weight, infu_df$Tumor_Volume_mm3), 2)), x = 25, y = 95)
```


```{r}
pairs(infu_df[,2:5], pch = 19, col = "green")
```

### ketapril_df:

```{r}
keta_df1 <- select(ketapril_df, Mouse_Id, Timepoint, Tumor_Volume_mm3) %>%
  group_by(Mouse_Id) %>%
  filter(Timepoint == max(Timepoint, na.rm=TRUE))

### Find the average weight by mice_id in Infubinol_df


keta_df2 <- select(ketapril_df, Mouse_Id, Weight_g) %>%
  group_by(Mouse_Id) %>%
 summarise(Average_weight = mean(Weight_g, na.rm=TRUE))

### Joining the two df's for adding average weight

keta_df <- keta_df1 %>% inner_join(keta_df2, by = "Mouse_Id")

### Find the average age by mice_id in Capomulin_df

keta_df3 <- select(ketapril_df, Mouse_Id, Age_months) %>%
  group_by(Mouse_Id) %>%
 summarise(Average_age = mean(Age_months, na.rm=TRUE))

### Joining the two df's for adding average age

keta_df <- keta_df %>% inner_join(keta_df3, by = "Mouse_Id")

head(keta_df)
```

### summerize the Tumor_Volume_mm3

```{r}

keta_df$Tumor_Volume_mm3 %>%
  summary()
```

### Standard Deviation

```{r}

keta_df$Tumor_Volume_mm3 %>% sd()
```

### ketapril_df Vs Age_months

```{r}

# Creating the plot
plot(keta_df$Average_age, keta_df$Tumor_Volume_mm3, pch = 19, col = "purple")

# Regression line
abline(lm(keta_df$Tumor_Volume_mm3 ~ keta_df$Average_age), col = "red", lwd = 3)

# Pearson correlation
text(paste("Correlation:", round(cor(keta_df$Average_age, keta_df$Tumor_Volume_mm3), 2)), x = 25, y = 95)
```

### ketapril_df Vs Weight_g

```{r}

# Creating the plot
plot(keta_df$Average_weight, keta_df$Tumor_Volume_mm3, pch = 19, col = "purple")

# Regression line
abline(lm(keta_df$Tumor_Volume_mm3 ~ keta_df$Average_weight), col = "red", lwd = 3)

# Pearson correlation
text(paste("Correlation:", round(cor(keta_df$Average_weight, keta_df$Tumor_Volume_mm3), 2)), x = 25, y = 95)
```

```{r}
pairs(keta_df[,2:5], pch = 19, col = "purple")
```


### placebo_df:

```{r}
plac_df1 <- select(placebo_df, Mouse_Id, Timepoint, Tumor_Volume_mm3) %>%
  group_by(Mouse_Id) %>%
  filter(Timepoint == max(Timepoint, na.rm=TRUE))

### Find the average weight by mice_id in Infubinol_df


plac_df2 <- select(placebo_df, Mouse_Id, Weight_g) %>%
  group_by(Mouse_Id) %>%
 summarise(Average_weight = mean(Weight_g, na.rm=TRUE))

### Joining the two df's for adding average weight

plac_df <- plac_df1 %>% inner_join(plac_df2, by = "Mouse_Id")

### Find the average age by mice_id in Capomulin_df

plac_df3 <- select(placebo_df, Mouse_Id, Age_months) %>%
  group_by(Mouse_Id) %>%
 summarise(Average_age = mean(Age_months, na.rm=TRUE))

### Joining the two df's for adding average age

plac_df <- plac_df %>% inner_join(plac_df3, by = "Mouse_Id")

head(plac_df)
```

### summerize the Tumor_Volume_mm3

```{r}

plac_df$Tumor_Volume_mm3 %>%
  summary()
```

### Standard Deviation

```{r}

plac_df$Tumor_Volume_mm3 %>% sd()
```


### placebo_df Vs Age_months

```{r}

# Creating the plot
plot(plac_df$Average_age, plac_df$Tumor_Volume_mm3, pch = 19, col = "lightblue")

# Regression line
abline(lm(plac_df$Tumor_Volume_mm3 ~ plac_df$Average_age), col = "red", lwd = 3)

# Pearson correlation
text(paste("Correlation:", round(cor(plac_df$Average_age, plac_df$Tumor_Volume_mm3), 2)), x = 25, y = 95)
```

### placebo_df Vs Weight_g

```{r}

# Creating the plot
plot(plac_df$Average_weight, plac_df$Tumor_Volume_mm3, pch = 19, col = "lightblue")

# Regression line
abline(lm(plac_df$Tumor_Volume_mm3 ~ plac_df$Average_weight), col = "red", lwd = 3)

# Pearson correlation
text(paste("Correlation:", round(cor(plac_df$Average_weight, plac_df$Tumor_Volume_mm3), 2)), x = 25, y = 95)
```

```{r}
pairs(plac_df[,2:5], pch = 19, col = "lightblue")
```

### Conclusion:

From the plots above, there seems a correlation between weight and Tumor size for capomulin drug regimen but will be checked by calculating the correlation coefficient.



Hypothesis Testing for Capomulin drug regimen:

sample size(n) = 25, sample Tumor size mean(xbar) =  36.67 mm3, standard devation  = 5.71 mm3

### Step 1: Formulate Hypothesis

Null hypothesis: There is no difference between the effectiveness of the four drug regimens. In other words, the mean of the size of Tumor for Capomulin drug regimen result is the same as Placebo. 

Null Hypothesis H^0: mu = 60.51
Alternate Hypothesis: mu != 60.51

It is a Two Tailed test. The claim is that the population mean is equal to mm3.
So you reject the claim if the sample mean is way above 60.51mm3 or way below 60.51mm3

### Step 2: Calculate the t_statistics

Here you translate this test on to t_ distribution by calculating the t_statistics

t_statistic = xbar-mu/s/square root of n


### loading the following required packages:
```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
```

### Two-sample t-test
The two-sample t-test is also known as the independent t-test. The independent samples t-test comes in two different forms:

the standard Student’s t-test, which assumes that the variance of the two groups are equal.
the Welch’s t-test, which is less restrictive compared to the original Student’s test. This is the test where you do not assume that the variance is the same in the two groups, which results in the fractional degrees of freedom.

### Preparing Data for calculating t-statistics

```{r}

capo_df1 <- select(capomulin_df, Mouse_Id, Drug_Regimen, Timepoint, Tumor_Volume_mm3) %>%
  group_by(Mouse_Id) %>%
  filter(Timepoint == max(Timepoint, na.rm=TRUE))

head(capo_df1)
```
### To select the specific columns capo_df1

```{r}
capo_df1 <- select(capomulin_df, Mouse_Id, Drug_Regimen, Tumor_Volume_mm3)

head(capo_df1)
```

```{r}

plac_df1 <- select(placebo_df, Mouse_Id, Timepoint, Drug_Regimen, Tumor_Volume_mm3) %>%
  group_by(Mouse_Id) %>%
  filter(Timepoint == max(Timepoint, na.rm=TRUE))

head(plac_df1)
```

```{r}
plac_df1 <- select(placebo_df, Mouse_Id, Drug_Regimen, Tumor_Volume_mm3) %>%
  group_by(Mouse_Id)

head(plac_df1)
```
### To select the specific columns plac_df1

```{r}

plac_df1 <- select(placebo_df, Drug_Regimen, Tumor_Volume_mm3)

plac_df1

```

```{r}
capo_df1 %>% full_join(plac_df1)
```

